---
title: "HW04"
author: "VY"
date: '`r Sys.Date()`'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Matrix)
library(microbenchmark)
library(parallel)
```

# Part 1: HPC

```{r rewrite functions}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  Matrix::rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

# interestingly, this method took a lot longer
# cl <- makePSOCKcluster(4) 
# x <-20
# clusterSetRNGStream(cl, 123) 
# clusterExport(cl, "x")
# fun2alt <- function(mat) {parLapply( cl= cl, X = mat, fun = t(apply(mat, 1, cumsum)))}
  

fun2alt <- function(mat) {
t(apply(mat, 1, cumsum))
}


# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "ms", check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "ms", check = NULL
)
```

# Problem 2

```{r, simulate pi}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132

# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})



# 2. PREPARING THE CLUSTER
cl <- makePSOCKcluster(4) 
x <-4000


system.time({
clusterSetRNGStream(cl, 123) 
clusterExport(cl, "x")
  ans <- unlist(parLapply(
    cl = cl,
    X=x,
    fun = sim_pi,
    n=10000
  ))
  print(mean(ans))
  stopCluster(cl)
})
```

# Part 2: SQL
```{r initialize connection}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

Question 1
How many many movies is there avaliable in each rating catagory.
```{sql connection=con}
SELECT rating, count(film_id) as count 
FROM film 
GROUP BY rating
```


Question 2
What is the average replacement cost and rental rate for each rating category.
```{sql connection=con}
SELECT rating, AVG(replacement_cost), AVG(rental_rate)
FROM film
GROUP BY rating
```

PG-13 movies had the highest average replacement cost, but PG movies had the highest average rental rate.

Question 3
Use table film_category together with film to find the how many films there are with each category ID
```{sql connection=con}
SELECT b.category_id, count(*) as count
FROM film AS a
  INNER JOIN film_category AS b
  ON a.film_id = b.film_id
GROUP BY b.category_id
```


Question 4
Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{sql connection=con}
SELECT b.category_id, c.name, count(*) as count
FROM film AS a
  INNER JOIN film_category AS b
  ON a.film_id = b.film_id
  INNER JOIN category AS c
  ON b.category_id = c.category_id
GROUP BY b.category_id
ORDER BY count DESC;
```

The most common category of movie in this sample was Sports.

```{r disconnect}
dbDisconnect(con)
```

