---
title: "HW01: PM2.5 concentration in California 2004 vs 2019"
author: "VY"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, force = TRUE)
webshot::install_phantomjs()
library(leaflet)

library(webshot)

library(dplyr)

library(data.table)

library(tidyverse)
```

# HW02
## Data Wrangling
```{r read in data and merge, cache=TRUE}
indchs <- read.csv(file = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv")
regchs <- read.csv(file = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv")
chs <- merge(indchs, regchs, by.x = "townname", by.y = "townname")
chs <- data.table(chs)
```

## 1. After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.
In the case of missing values, impute data using the average within the variables “male” and “hispanic.” If you are interested (and feel adventurous) in the theme of Data Imputation, take a look at this paper on “Multiple Imputation” using the Amelia R package here.

## The individual dataset had `r nrow(indchs)` rows and the regional dataset had `r nrow(regchs)` rows. The merged dataset has `r nrow(chs)` rows, so no data was lost.
```{r data imputation, cache=TRUE}
summary(chs)
# agepft height weight bmi asthma father_asthma mother_asthma wheeze hayfever allergy educ_parent smoke gasstove fev fvc mmef no_24hr pm2_5_fr have missing data
hismalemean <- chs[hispanic==1 & male==1]
hismaleval <- rbind(agepftm = mean(hismalemean$agepft, na.rm=T), heightm = mean(hismalemean$height, na.rm=T), weightm =mean(hismalemean$weight, na.rm=T), bmim=mean(hismalemean$bmi, na.rm=T), asthmam =mean(hismalemean$asthma, na.rm=T), father_asthmam =mean(hismalemean$father_asthma, na.rm=T), mother_asthmam =mean(hismalemean$mother_asthma, na.rm=T), wheezem =mean(hismalemean$wheeze, na.rm=T), hayfeverm =mean(hismalemean$hayfever, na.rm=T), allergym =mean(hismalemean$allergy, na.rm=T), educ_parentm =mean(hismalemean$educ_parent, na.rm=T), smokem =mean(hismalemean$smoke, na.rm=T), gasstovem =mean(hismalemean$gasstove, na.rm=T), fevm =mean(hismalemean$fev, na.rm=T), fvcm =mean(hismalemean$fvc, na.rm=T), mmefm =mean(hismalemean$mmef, na.rm=T), no_24hrm =mean(hismalemean$no_24hr, na.rm=T), pm2_5_frm =mean(hismalemean$pm2_5_fr, na.rm=T))
chs1 <- chs[ , agepft := replace(agepft, is.na(agepft), 9.966942e+00)]
chs1 <- chs[,          height := replace(height, is.na(height), 1.385984e+02)]
chs1 <- chs[,  weight := replace(weight, is.na(weight), 8.276707e+01)]
chs1 <- chs[, bmi := replace(bmi, is.na(bmi), 1.941148e+01)]
chs1 <- chs[, asthma := replace(asthma, is.na(asthma), 1.601562e-01)]
chs1 <- chs[, father_asthma := replace(father_asthma, is.na(father_asthma), 8.403361e-02)]
chs1 <- chs[,  mother_asthma := replace(mother_asthma, is.na(mother_asthma), 1.067194e-01)]
chs1 <- chs[,  wheeze := replace(wheeze, is.na(wheeze), 3.534137e-01)]
chs1 <- chs[,  hayfever := replace(hayfever, is.na(hayfever), 1.744681e-01)]
chs1 <- chs[, allergy := replace(allergy, is.na(allergy), 2.540323e-01)]
chs1 <- chs[,  educ_parent := replace(educ_parent, is.na(educ_parent), 2.423868e+00)]
chs1 <- chs[, smoke := replace(smoke, is.na(smoke), 1.501976e-01)]
chs1 <- chs[, gasstove := replace(gasstove, is.na(gasstove), 8.156863e-01)]
chs1 <- chs[,  fev := replace(fev, is.na(fev), 2.120266e+03)]
chs1 <- chs[, fvc := replace(fvc, is.na(fvc), 2.443876e+03)]
chs1 <- chs[, mmef := replace(mmef, is.na(mmef), 2.447494e+03)]
chs1 <- chs[, no_24hr := replace(no_24hr, is.na(no_24hr), 1.586585e+01)]
chs1 <- chs[, pm2_5_fr := replace(pm2_5_fr, is.na(pm2_5_fr), 1.994291e+01)]
      
```

The original dataset had `r sum(is.na(chs))` missing values. After replacing NAs, there are `r sum(is.na(chs1))` missing values.

## 2. Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.
```{r obesity level}
chs1 <- mutate(chs1, obesity_level = ifelse(chs1$bmi<14, "underweight", ifelse(14<=chs1$bmi & chs1$bmi < 22, "normal", ifelse(22<=chs1$bmi & chs1$bmi < 24, "overweight", ifelse(chs1$bmi>=24, "obese", NA)))))
chs1 %>% group_by(obesity_level) %>% summarize(mean_bmi = mean(bmi), min_bmi = min(bmi), max_bmi = max(bmi), n_bmi = n()) %>%
  knitr::kable()
```
## 3. Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.
```{r}
chs1 <- mutate(chs1, smoke_gas_exposure = ifelse(smoke==1 & gasstove==1, "both", ifelse(smoke==1, "smoke", ifelse(gasstove==1, "gasstove", "none"))))
table(chs1$smoke_gas_exposure)
```

## 4. Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”
```{r summary table}
townsum <- chs1 %>% group_by(townname) %>% summarise(mean_fev = mean(fev), sd_fev = sd(fev), freq_asthma = sum(asthma)/ n()) %>% distinct(townname, .keep_all = TRUE) 
townsum%>% knitr::kable()

sexsum <- chs1 %>% group_by(male) %>% summarise(mean_fev = mean(fev), sd_fev = sd(fev), freq_asthma = sum(asthma)/ n()) %>% distinct(male, .keep_all = TRUE) 
sexsum%>% knitr::kable()

obesitysum <- chs1 %>% group_by(obesity_level) %>% summarise(mean_fev = mean(fev), sd_fev = sd(fev), freq_asthma = sum(asthma)/ n()) %>% distinct(obesity_level, .keep_all = TRUE) 
obesitysum%>% knitr::kable()

smokegassum <- chs1 %>% group_by(smoke_gas_exposure) %>% summarise(mean_fev = mean(fev), sd_fev = sd(fev), freq_asthma = sum(asthma)/ n()) %>% distinct(smoke_gas_exposure, .keep_all = TRUE) 
smokegassum %>% knitr::kable()

```

# Exploratory Data Analysis
## The primary questions of interest are: 1. What is the association between BMI and FEV (forced expiratory volume)? 2. What is the association between smoke and gas exposure and FEV? 3. What is the association between PM2.5 exposure and FEV?
## Check for any implausible values
```{r check implausible values}
summary(chs1$bmi)
summary(chs1$fev)
summary(chs1$pm2_5_fr)
histobmi <- ggplot(chs1, aes(x=bmi)) +
  geom_histogram(binwidth=0.5, fill="blue", position="dodge") +
  theme(legend.position="right")
histobmi
histofev <- ggplot(chs1, aes(x=fev)) +
  geom_histogram(binwidth=10, fill="red", position="dodge") +
  theme(legend.position="right")
histofev
histopm2.5 <- chs1 %>% ggplot(aes(x=pm2_5_fr)) +
  geom_histogram(binwidth=2, fill="green", position="dodge") +
  theme(legend.position="right")
histopm2.5
```

There is a peak bmi at 19.41, peak fev at 2120.3, and peak PM2.5 at 19.9 because the missing values were assigned the average value of a Hispanic male.

## 1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
```{r scatterplots with regression lines}
chs1 %>% ggplot(aes(x=bmi, y=fev, group=townname, color=townname)) +
  stat_smooth(method = "lm", formula = y ~ x, geom="smooth") +
  geom_point(shape= ".") +
  theme(legend.position="right") +
  facet_wrap(vars(townname), nrow = 4) +
  ggtitle("BMI vs FEV1 by town") +
  xlab("BMI") +
  ylab("FEV1")
```

Interpretation: BMI and FEV1 appear to have a positive linear relationship, regardless of town. The towns where the positive linear relationship looks the most significant are San Dias and Lake Elsinore.

## 2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r stacked histograms}
fevobhisto <- chs1 %>% ggplot(aes(x=fev, group=obesity_level, fill=obesity_level)) +
  geom_histogram(position = "stack", binwidth = 10) +
  scale_fill_manual(values = c("cyan3", "orange", "yellow", "deeppink")) +
  ggtitle("Distribution of FEV1 by BMI category") +
  labs(fill= "BMI Category")
fevobhisto
fevsmg <- chs1 %>% ggplot(aes(x=fev, group=smoke_gas_exposure, fill=smoke_gas_exposure)) +
  geom_histogram(position = "stack", binwidth = 20) +
  scale_fill_manual(values = c("coral", "bisque", "chocolate", "darksalmon")) +
  ggtitle("Distribution of FEV1 by Smoke or Gas exposure") +
  labs(fill= "Presence of smoke/gas/both exposure")
fevsmg
```

Interpretation: There is a clear mode at 2120.3, the value of the imputed data. Otherwise, it appears that presence of  smoke or gas or both exposure do not have significantly different distributions of FEV1.

## 3. Barchart of BMI by smoke/gas exposure.
```{r barchart bmi smoke/gas}
bmismg <- chs1 %>% ggplot(aes(x=bmi, group=smoke_gas_exposure, fill=smoke_gas_exposure)) +
  geom_histogram(position = "stack", binwidth = 0.2) +
  ggtitle("Distribution of BMI by Smoke or Gas exposure") +
  labs(fill= "Presence of smoke/gas/both exposure")
bmismg
```

Interpretation: Again, the imputed mean Hispanic male BMI is a clear mode. It appears that the distribution of bmi in all smoke/gas exposure categories appears similar. However, the tail appears to be of only subjects with gasstove or both exposure.

## 4. Statistical summary graphs of FEV by BMI category and FEV by smoke/gas exposure category. 
```{r FEVbmi FEVsge}
bmiavg <- chs1[ , .(mean_fev = mean(fev)), by = "obesity_level"]
bmiavg %>% knitr::kable()
chs1 %>% ggplot(mapping = aes(x = obesity_level, y = fev)) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange") +
  stat_summary(fun.data = mean_sdl, geom = "errorbar")

sgeavg <- chs1[ , .(mean_fev = mean(fev)), by = "smoke_gas_exposure"]
sgeavg %>% knitr::kable()
chs1 %>% ggplot(mapping = aes(x = smoke_gas_exposure, y = fev)) +
  stat_summary(fun.data = mean_sdl, geom = "pointrange") +
  stat_summary(fun.data = mean_sdl, geom = "errorbar")
```

Interpretation: 
Obese and overweight children appear to have higher FEV1, where as normal and underweight BMI categories have lower FEV1.

There does not appear to be a relationship between smoke/gas exposure and FEV1. The average FEV1 by exposure appears similar in value, with smoke exposure only having a slightly higher FEV1.

## 5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
```{r leaflet}

pm.pal <- colorFactor(topo.colors(12), domain=chs1$townname)
pm.pal

pmmap <- chs1 %>% group_by(townname) %>% mutate(mean_pm2.5 = mean(pm2_5_fr)) %>% distinct(townname, .keep_all = TRUE) %>% leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addCircles(
    lat=~lat, lng=~lon,
    label=~paste0(mean_pm2.5), labelOptions = labelOptions(noHide = T, textOnly = TRUE), color = ~pm.pal(townname),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  addLegend("bottomright", pal=pm.pal, values= chs1$townname, title = "Town") %>%
  addControl("Average PM2.5 levels by Town", position = "topright")
pmmap
``` 

Interpretation: The highest mean PM2.5 levels appear to be clustered around LA, especially NE of LA.

## 6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.
```{r PM2.5 FEV}
chs1 %>% ggplot(aes(x=pm2_5_fr, y=fev)) +
  stat_smooth(method = "lm", formula = y ~ x, geom="smooth") +
  geom_point(shape= ".") +
  theme(legend.position="right") +
  ggtitle("PM2.5 vs FEV1") +
  xlab("PM2.5") +
  ylab("FEV1")
```

Interpretation: There appears to be a slightly negative relationship between FEV1 and PM2.5, where higher levels of PM2.5 are associated with slightly lower average FEV1.




