---
title: "HW03"
author: "VY"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(httr)
library(tidyverse)
library(stringr)
```

## Data Scraping

Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r find the papers}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")
# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")
# Turning it into text
counts <- as.character(counts)
# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```


Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r get the papers}
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db      = "pubmed",
    term    = "sars-cov-2 trial vaccine",
    retmax  = 250
  ), 
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)

ids <- as.character(ids)
# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]
# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")
```


As we did in lab 7. Create a dataset containing the following:

Pubmed ID number,
Title of the paper,
Name of the journal where it was published,
Publication date, and
Abstract of the paper (if any).

```{r create abstract dataset}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db  = "pubmed",
    id  = paste(ids,collapse = ","),
  retmax = 250,
  rettype = "abstract"
    )
)
# Extracting the content of the response of GET
publications <- httr::content(publications)

pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)

# Get abstracts
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")

abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>") 

abstracts <- str_replace_all(abstracts, "[[:space:]]+"," ")


# Get titles
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")

titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")

# Get journal
journals <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")

journals <- str_remove_all(journals, "</?[[:alnum:]- =\"]+>")

# Get publication date
dates <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")

dates <- str_remove_all(dates, "</?[[:alnum:]- =\"]+>\n")

# Dataset
database <- data.frame(
  PubMedId = ids,
  Title    = titles,
  Journal = journals,
  PubDate = dates,
  Abstract = abstracts
)

knitr::kable(database[1:5,], caption = "Some papers about Covid19 trial vaccine")
```

