---
title: "HW03"
author: "VY"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(httr)
library(tidyverse)
library(tidytext)
library(stringr)
library(data.table)
library(dplyr)
library(ggplot2)
library(forcats)
```

# Data Scraping

## Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r find the papers}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")
# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")
# Turning it into text
counts <- as.character(counts)
# Extracting the data using regex
# stringr::str_extract(counts, "[0-9,]+")
```

There are `r stringr::str_extract(counts, "[0-9,]+")` papers on "sars-cov-2 trial vaccine".

## Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r get the papers}
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db      = "pubmed",
    term    = "sars-cov-2 trial vaccine",
    retmax  = 250
  ), 
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)

ids <- as.character(ids)
# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]
# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")
```


## As we did in lab 7. Create a dataset containing the following:

## Pubmed ID number, Title of the paper, Name of the journal where it was published, Publication date, and Abstract of the paper (if any).

```{r create abstract dataset}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db  = "pubmed",
    id  = paste(ids,collapse = ","),
  retmax = 250,
  rettype = "abstract"
    )
)
# Extracting the content of the response of GET
publications <- httr::content(publications)

pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)

# Get abstracts
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")

abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>") 

abstracts <- str_replace_all(abstracts, "[[:space:]]+"," ")


# Get titles
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")

titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")

# Get journal
journals <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")

journals <- str_remove_all(journals, "</?[[:alnum:]- =\"]+>")

# Get publication date
dates <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")

dates <- str_remove_all(dates, "</?[[:alnum:]- =\"]+>\n")

# Dataset
database <- data.frame(
  PubMedId = ids,
  Title    = titles,
  Journal = journals,
  PubDate = dates,
  Abstract = abstracts
)

knitr::kable(database[1:5,], caption = "Some recent papers about Covid19 trial vaccine")
```


# Text Mining
## A new dataset has been added to the data science data repository https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed. The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.

```{r read in data}
pubmed <- read.csv(file = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv")

pubmed <- data.table(pubmed)
```


## Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

```{r tokenize abstracts}
pubmed %>%   group_by(term) %>%
  unnest_tokens(word, abstract) %>%
  count(word, sort = TRUE) %>% 
  arrange(desc(n)) %>%
  group_by(term) %>%
  slice(1:5) %>%
  knitr::kable(caption = "Most common tokens for each search term with stopwords")

```

The top words, without removing stop words, are largely uninteresting. Stop words occupy the top 5 most frequent words in the data set for each of the search terms except for prostate cancer, which returned "cancer" and "prostate" at 4th and 5th place.

```{r remove stopwords}
pubmed %>%   group_by(term) %>%
  unnest_tokens(word, abstract) %>%
  anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>% 
  arrange(desc(n)) %>%
  group_by(term) %>%
  slice(1:5) %>%
  knitr::kable(caption = "Most common tokens for each search term WITHOUT stopwords")
```

After removing the stop words, the top 5 words for each term are much different and appear to describe the term word better. It is interesting that "patient" appears in the top 5 for each term except for preeclampsia, where "women" is in the top 5 instead.

## Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.

```{r create bigrams}
pubmed %>%
  unnest_ngrams(bigram, abstract, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(n, fct_reorder(bigram, n))) +
  geom_col() +
  ylab("Abstract Bigram")+
  ggtitle("Top 10 most common bigrams in the abstract dataset")
```

Some stopwords are top bigrams(of the, in the, and the, to the), however covid 19 is the most common bigram in the abstracts. Some of the other search terms appear here as well (pre eclampsia, prostate cancer)

## Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r tf idf values}
pubmed %>%
  unnest_tokens(word, abstract) %>%
  count(word, term) %>%
  bind_tf_idf(word, term, n) %>%
  arrange(desc(tf_idf)) %>%
  group_by(term) %>%
  slice(1:5) %>%
  knitr::kable(caption = "Highest TF-IDF value words by search term")
```

Looking at TF-IDF values, the highest value word is the same as the most common word, except for cystic fibrosis, where the highest TF-IDF word is "cf" whereas the most common word is "fibrosis". The rest of the top 5 words per term appear to be more descriptive of the term than the most common words (ie for term "meningitis", we see "pachymeningitis" as a top TF-IDF word whereas we saw "clinical" and "patient" as most common words).